---
output: github_document
bibliography: Vignettes/literature.bib 
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## EpisodeR ##

<!-- badges: start -->
<!-- badges: end -->

An R package to load, explore, and work with any index variable in the V-Dem data set - a project of the [V-Dem Institute](https://www.v-dem.net/) - to built episodes of growth and decline in a specific V-Dem index variable. This package was used for the Academic Freedom Growth and Decline Episodes data set introduced in @lott_academic_2023.

This package uses code parts of the [ERT package](https://github.com/vdeminstitute/ERT) (@institute_vdeminstituteert_2023). All copied and adapted code parts are marked accordingly in the R code of this package. However, unlike the [ERT package](https://github.com/vdeminstitute/ERT), this package uses a different approach to identify episodes of change in a variable. These differences are documented below. 

#### Acknowledgements ####

I would like to thank the authors of the [ERT package](https://github.com/vdeminstitute/ERT), namely Seraphine Maerz, Amanda Edgell, Joshua Krusell, Laura Maxwell, and Sebastian Hellmeier, who have licensed the [ERT package](https://github.com/vdeminstitute/ERT) under the GPL-3 license. This package (EpisodeR) is also licensed under the GPL-3 license.  Without the excellent documentation and work of these colleagues, this package would not have been possible. 

The work of the Episodes of Regime Transformation team has published as @maerz_episodes_2023, @edgell_institutional_2022,  @wilson_episodes_2023, @boese_how_2021, among others. The package development has been funded by a Volkswagen Foundation [grant number A138109, PI: Katrin Kinzelbach and Staffan I. Lindberg] and is part of the [Academic Freedom Index](https://academic-freedom-index.net/) project. 

## EpisodeR package ##

#### Load, explore, and work the episodes data sets ####
* Note: for non-R users I provide the *Academic Freedom Growth and Decline Episodes data set* in this GitHub project as [.csv](https://github.com/LarsLott/EpisodeR/tree/master/inst) files. 
* RELEASE: Academic Freedom Growth and Decline Episodes data set 15.0 is based on the V-Dem dataset v15. In the following years, updated *Academic Freedom Growth and Decline Episodes data set* will be provided once a year when updated V-Dem is published. 

#### Functions ####
* `get_episode`: Identify episodes of growth and decline in an index variable in which this index variable systematically grows or declines in respective and connected country-years. This functions controls for overlapping uncertainty intervals, "var_codelow" and "var_codehigh", as suggested by @pelke_conceptualizing_2021. A growth episode in an index variable is defined as a cumulative
increase of 0.1 or more on any index variable. A decline episode is defined as a a cumulative drop of 0.1 or more on any index variable. 
* `get_episode_wo_CI`: Identify episodes of growth and decline in an index variable in which this index variable systematically grows or declines in respective and connected country-years. This functions **does not** control for overlapping uncertainty intervals, "var_codelow" and "var_codehigh", as suggested by @pelke_conceptualizing_2021. A growth episode in an index variable is defined as a cumulative increase of 0.1 or more on any index variable. A decline episode is defined as a a cumulative drop of 0.1 or more on any index variable.
* `plot_episodes`: Plot growth and decline episodes over time for a selected country.
* `plot_all_episodes`: Plot share or absolute number of all countries in growth and decline episodes of a specific index variable over time.

## Differences between ERT package and EpisodeR package ##

The [ERT package](https://github.com/vdeminstitute/ERT) differs from this package in important ways. The ERT package is computationally more efficient and more flexible in what constitutes an episode of change in a variable. It was one important source of code for this package it inspired my work for this package. However, two drawbacks come with the [ERT package](https://github.com/vdeminstitute/ERT): 

* It does not enable users to search for episode of regime transformation / change other than the Electoral Democracy Index. 
* Secondly, the ERT package does not control for overlapping uncertainty intervals (before the start of an episode and at the end of an episode), as suggested by @pelke_conceptualizing_2021. 

These drawbacks are partially resolved in this package, while it comes with other important drawbacks: It is computationally inefficient and less flexible in terms of episode termination and tolerance for temporary stagnation. 

The `EpisodeR` package copied and adapted some code parts of the ERT package but does not use the most important part of the ERT package (C++ code for finding episodes) to estimate episodes of change data. It differs most substantially here: 

* The ERT package uses C++ language to find episodes of regime transformation, while this package uses some loops in R for finding these episodes of change in an index variable. The `EpisodeR` package is computationally more inefficient compared to the C++ solution from the ERT package. To find episodes of change in a variable, this package uses code originally used by @pelke_conceptualizing_2021. 
* The ERT package enables users to consider a specific number of years as tolerance for stasis or a gradual movement in the opposite direction. By default it is set to **five years**. In the `EpisodeR` package users cannot change the tolerance parameter and there is a temporary stasis on the specific variable with no further increase/decline of `start_incl` points in **four years**. 

## Installation ##

You can install the development version of EpisodeR from [GitHub](https://github.com/larslott/EpisodeR) with:

``` r
# Install the development version of the EpisodeR package 
# (since this package is still an ongoing project, 
# keep checking for updates, new functions, etc.!)

# First, you need to have the devtools package installed
install.packages("devtools")
# now, install the EpisodeR package directly from GitHub
devtools::install_github("https://github.com/larslott/EpisodeR")

# installed. If you have troubles with the installation 
# write to the package maintainer Lars Lott (lars.lott@fau.de).

```

## Example ##

This is a basic example which shows you how to use the **EpisodeR** package:

```{r }
library(EpisodeR)
## basic example code
```

You can use `get_episode` to return a data.frame identifying decline and growth episodes of an index variable in which this variable systematically grow or decline in respective and connected country-years.  

```{r example}
df <- get_episode(data = EpisodeR::vdem,
                  start_incl = 0.01,
                  cum_incl = 0.1,
                  year_turn = 0.03,
                  variable = "v2xca_academ")
```

Vignettes explaining the different functions and how they differ from the ERT package are available in the Vignettes folder in the GitHub repository.  

## References ##

