---
output: github_document
bibliography: literature.bib 
---

# plot_episodes #


<!-- plot_episodes.md is generated from plot_episodes.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/plot_episode-",
  out.width = "100%"
)
```


In the context of this package, the `plot_episodes()` function returns a ggplot object and produces a plot that shows
growth and decline episodes for a selected country over time for a selected variable. 
The `plot_episodes()` function does not control for overlapping uncertainty intervals before the start of an episode an at the end of an episode. 

## The `plot_episodes()` function

By default, `plot_episodes()` returns a ggplot() object in R implementing the episode approach to measure growth and decline episodes of the Academic Freedom Index as suggested in @lott_academic_2023: 

```{r}
library(EpisodeR)

fig1 <- plot_episodes()
fig1
```

The `plot_episodes` function does not consider measurement uncertainty in the measurement of the index variables. 

By using the default values, users get a ggplot() object comparable to Figures 3 and 5 presented in @lott_academic_2023. However, users are also able to customize the default parameters and use other V-Dem index variables than the Academic Freedom Index. 

## Customizing the `plot_episodes()` function with user-specific parameters and variables ##

There are different ways users can customize the parameters and set other V-Dem index variables. Users can use the following arguments: `years`, `country`, `variable`, `var_label`, `start_incl`, `cum_incl`, `year_turn`,

### `years`: Vector with two numeric values indicating the minimum and maximum year to be plotted. ###

Vector with two numeric values indicating the minimum and maximum year to be plotted. By default, the vector is set to c(1900, 2022). 

```{r}
fig2 <- plot_episodes(years = c(1945, 2022))
fig2
```

### `country`: Character vector containing the country for which episodes should be shown ###

Only entries from the country_name column in the V-Dem dataset are accepted. The default country is "Sweden". 

```{r}
fig3 <- plot_episodes(country = c("India"))
fig3
```

### `variable`: Change the index variable that constitutes episodes of change ###

The `variable` argument enable users to get episodes for any V-Dem index variable, e.g. the Liberal Democracy Index (LDI), the Vertical Accountability Index, and the Liberal Component Index. Users should be extremely cautions in using variables, which are not scaled between 0 and 1, even it is possible to use other variable from the V-Dem universe. By default this variable is set to "v2xca_academ".   

### `var_label`:  What is the variable label that should be plotted? ###

By default, the label is set to "Academic Freedom Index". The label should correspond to the `variable` chosen.


```{r}
fig4 <- plot_episodes(variable = "v2x_libdem", var_label = "Liberal Democracy Index", country = "India")
fig4
```

### `start_incl`: Set the paramater that is necessary to trigger the start of an episode ###

The `start_incl` argument enable users to change the parameter that is necessary to trigger the start of an episode. This is the absolute value of the first difference. By default, this value is set to 0.01. The `plot_episodes()` function follows the potential episode as long as there is continued increase/decline, while allowing up to four years of temporary stagnation, meaning no further increase/decline of `start_incl` points or more on the respective variable.

### `cum_incl`: Set the minimum amount of total change on the index variable necessary to constitute a growth or decline episode ###

The `cum_incl` argument allows for changing the minimum amount of total change on the index variable necessary to constitute a growth or decline episode? To identify substantive growth and decline episodes, `plot_episodes()` function calculates the total magnitude of change from the year before the start of an episode to the end of an episode. This cumulative increase/drop is set to 0.1 by default (10% of the total 0–1 scale). It thus records only those manifest growth episodes which add up to a positive change of at least  `cum_incl` and as manifest decline episodes only those which add up to a negative change of at least `cum_incl`.

### `year_turn`: Set the parameter that is necessary to trigger the end of an episode ###

The `year_turn` argument enable users to change the parameter that is necessary to trigger the end of an episode. By default, this value is set to 0.03. The `plot_episodes()` function terminates an episode when there is a temporary stagnation on the AFI with no further increase/decline of `start_incl` points in four years or when the AFI decreases/increases by `year_turn` points, from one year to the next. 

## References ##
